### 2024-07-22 mythbusters.qmd
20240722 - fix citation

Considerable speculation exists regarding whether players perform better in their last year of their contract (i.e., their "contract year").
Fantasy football talking heads and commentators frequently discuss the benefit of selecting players who are in their contract year, because it supposedly means that player has more motivation to perform well so they get a new contract and get paid more.
To our knowledge, no peer-reviewed studies have examined this question for football players.
One study found that National Basketball Association (NBA) players improved in field goal percentage, points, and player efficiency rating (but not other statistics: rebounds, assists, steals, or blocks) from their pre-contract year to their contract year, and that Major League Baseball (MLB) players improved in runs batted in (RBIs; but not other statistics: batting average, slugging percentage, on base percentage, home runs, fielding percentage) from their pre-contract year to their contract year [White2014a].
Other casual analyses have been examined contract-year performance of National Football League (NFL) players, including articles in [2012](https://www.4for4.com/2012/preseason/2012-contract-year-players-and-myth-increased-production) (archived [here](https://perma.cc/CT3F-QN5E)) and [2022](https://www.4for4.com/2022/preseason/do-players-perform-better-fantasy-football-contract-year) (archived [here](https://perma.cc/F4F5-7RQZ)).

>>>>

Considerable speculation exists regarding whether players perform better in their last year of their contract (i.e., their "contract year").
Fantasy football talking heads and commentators frequently discuss the benefit of selecting players who are in their contract year, because it supposedly means that player has more motivation to perform well so they get a new contract and get paid more.
To our knowledge, no peer-reviewed studies have examined this question for football players.
One study found that National Basketball Association (NBA) players improved in field goal percentage, points, and player efficiency rating (but not other statistics: rebounds, assists, steals, or blocks) from their pre-contract year to their contract year, and that Major League Baseball (MLB) players improved in runs batted in (RBIs; but not other statistics: batting average, slugging percentage, on base percentage, home runs, fielding percentage) from their pre-contract year to their contract year [@White2014a].
Other casual analyses have been examined contract-year performance of National Football League (NFL) players, including articles in [2012](https://www.4for4.com/2012/preseason/2012-contract-year-players-and-myth-increased-production) (archived [here](https://perma.cc/CT3F-QN5E)) and [2022](https://www.4for4.com/2022/preseason/do-players-perform-better-fantasy-football-contract-year) (archived [here](https://perma.cc/F4F5-7RQZ)).
